SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-10-26 maintex.tex
Removed a couple of duplicate paragrphs

The MaNGA survey targets galaxies from a subset of 41,154 galaxies in the NASA-Sloan Atlas (NSA v1\_0\_1; \url{http://www.nsatlas.org}) with a redshift range of 0.01 $< z <$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_i}$ $<$ -24.0, where $\mathcal{M_i}$ is the rest frame $i$-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 \reff\ and 2.5 \reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the 8th MaNGA Product Launch (MPL-8), which covers 6142 unique galaxies observed by July 3, 2018.

@@

\begin{equation}\label{eq:blue}
NUV-r = 3.1682 - 0.16 (\mathcal{M}_i+18)
\end{equation}
\begin{equation}\label{eq:red}
NUV-r = 4.7866 - 0.04 (\mathcal{M}_i+18),
\end{equation}
Where $NUV-r$ is the color from SDSS's $k$-corrected absolute magnitude and $\mathcal{M}_i$ is the $i$-band magnitude from the NSA catalog. We further apply an \ewha\ $\ge$ 3\AA\ cut on galaxy spectra extracted from a 1.0~\reff\ aperture to ensure that all passive galaxies are removed from the sample \citep{Cid-Fernandes:2011}. Finally, we use the BPT diagnostic \citep{Baldwin:1981} to remove galaxies with possible AGN in their centers. Specifically, we use the emission line ratios, log([O~{\sc iii}]/H$\beta$) and log([N~{\sc ii}]/H$\alpha$), extracted from a 1.3~kpc aperture along with the maximum starburst line of \citet{Kewley:2001} as the demarcation between the star-forming branch and the AGN branch.

@@

From the NSA catalog's 641,409 galaxies, we find 492 galaxies which are paired to MaNGA targets. After restricting the sample to SFGs, we have 133 MaNGA targets with paired galaxies outside of the IFU. MaNGA targets which have both an inside-IFU and an outside-IFU pair are left to the inside-IFU sample. This gives us a total of 175 MaNGA targets in our pair sample.

@@

It has been shown that the SFR in galaxies is dependent on both the stellar mass and the redshift of the galaxies \citep{Noeske:2007}. We will compare our interacting galaxies with isolated galaxies of a similar stellar mass and redshift, to account for these other dependencies. To do this, we will use two different methods of pair - control comparison.

In the first method, we only control for the stellar mass between the pairs and controls. We split both the pair and control samples into five evenly spaced stellar mass bins over the range, \logm\ $=$ 9.0$-$11.5. The pairs are then compared to their respective controls within each stellar mass bin. While the redshift is not constrained in this method, we will show in Section \ref{sec:mass-bin} that this method is sufficient to reveal the merger induced star formation.

@@

We fit the geometry with apertures from both catalogs when available. When using the first method of pair-control comparison, where pairs and controls are grouped into stellar mass bins, we fit the control galaxies with the NSA apertures. When making the comparison with the second tailored control method, paired galaxies fitted with the NSA aperture are compared against controls fitted with the NSA aperture and paired galaxies fitted with the \citet{Simard:2011} apertures are compared against controls fitted with the \citet{Simard:2011} apertures. Further, when defining a galaxy's geometry with the \citet{Simard:2011} apertures we also use the masses given in the catalog for the total stellar mass of the galaxy.

We fit the geometry with apertures from both catalogs when available. When using the first method of pair-control comparison, where pairs and controls are grouped into stellar mass bins, we fit the control galaxies with the NSA apertures. When making the comparison with the second method, paired galaxies fitted with the NSA aperture are compared against controls fitted with the NSA aperture and paired galaxies fitted with the \citet{Simard:2011} apertures are compared against controls fitted with the \citet{Simard:2011} apertures. Further, when defining a galaxy's geometry with the \citet{Simard:2011} apertures we also use the masses given in the catalog for the total stellar mass of the galaxy. Note that although we carefully treat the two catalogs separately whenever possible, we find excellent agreement in the geometry parameters of isolated galaxies in both catalogs.

@@

Once the geometries of the galaxies are set, we can build azimuthally averaged radial profiles. The spaxels are binned into radius increments of 0.2 \reff\ from 0.0$-$2.6 \reff. While the MaNGA sample is limited spatially to 1.5$-$2.5 \reff\, the spaxels along the semi-minor axis of a galaxy will be able to extend beyond that range because of the inclination angle, sin($i$).

@@

Figure~\ref{fig:nuc_sep} compares the enhancement in both SFR and sSFR as a function of projected separation. For our data points, we use the mean $\Delta\log{\rm SFR}$ and $\Delta\log{\rm sSFR}$ measured within a deprojected radius of 0.5~\reff\ (i.e., the black data points in the right side panel of Figure~\ref{fig:ssfr_sep}). The literature data are either measured from single 1-1.5\arcsec-radius SDSS fibers \citep{Ellison:2008,Scudder:2012,Patton:2013} or XX\arcsec-radius apertures from CALIFA IFS data \citep{Bustamante:2020}. 
%The $\Delta$log(SFR) as a function of projected separation has been studied before with the SDSS survey \citep{Ellison:2008, Li:2008, Patton:2013, Scudder:2012, Bustamante:2020}. We compare the mentioned surveys with our own work in Figure \ref{fig:nuc_sep}. 
We find that our central SFR enhancements for closely separated pairs are higher than many of the previous studies, our sSFR enhancement is $\sim$0.2 dex higher than \citet{Scudder:2012} and $\sim$0.3 dex higher than \citet{Ellison:2008} and \citet{Bustamante:2020}. The central enhancement to the sSFR for closely separated pairs in our sample is $\sim$0.1 dex higher than \citet{Patton:2013}. The sSFR enhancement function from \citet{Li:2008} follows closely to our own and remains $\sim$0.05 dex above our sample out to 40 kpc.

>>>>

The MaNGA survey targets galaxies from a subset of 41,154 galaxies in the NASA-Sloan Atlas (NSA v1\_0\_1; \url{http://www.nsatlas.org}) with a redshift range of 0.01 $< z <$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M}_{\rm i}$ $<$ -24.0, where $\mathcal{M}_{\rm i}$ is the rest frame $i$-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 \reff\ and 2.5 \reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the 8th MaNGA Product Launch (MPL-8), which covers 6142 unique galaxies observed by July 3, 2018.

@@

\begin{equation}\label{eq:blue}
NUV-r = 3.1682 - 0.16 (\mathcal{M}_{\rm i}+18)
\end{equation}
\begin{equation}\label{eq:red}
NUV-r = 4.7866 - 0.04 (\mathcal{M}_{\rm i}+18),
\end{equation}
Where $NUV-r$ is the color from SDSS's $k$-corrected absolute magnitude and $\mathcal{M}_{\rm i}$ is the $i$-band magnitude from the NSA catalog. We further apply an \ewha\ $\ge$ 3\AA\ cut on galaxy spectra extracted from a 1.0~\reff\ aperture to ensure that all passive galaxies are removed from the sample \citep{Cid-Fernandes:2011}. Finally, we use the BPT diagnostic \citep{Baldwin:1981} to remove galaxies with possible AGN in their centers. Specifically, we use the emission line ratios, log([O~{\sc iii}]/H$\beta$) and log([N~{\sc ii}]/H$\alpha$), extracted from a 1.3~kpc aperture along with the maximum starburst line of \citet{Kewley:2001} as the demarcation between the star-forming branch and the AGN branch.

@@

@@

@@

We fit the geometry with apertures from both catalogs when available. When using the first method of pair-control comparison, where pairs and controls are grouped into stellar mass bins, we fit the control galaxies with the NSA apertures. When making the comparison with the second method, paired galaxies fitted with the NSA apertures are compared against controls fitted with the NSA apertures and paired galaxies fitted with the \citet{Simard:2011} apertures are compared against controls fitted with the \citet{Simard:2011} apertures. Further, when defining a galaxy's geometry with the \citet{Simard:2011} apertures we also use the masses given in the catalog for the total stellar mass of the galaxy. Note that although we carefully treat the two catalogs separately whenever possible, we find excellent agreement in the geometry parameters of isolated galaxies in both catalogs.

@@

Once the geometries of the galaxies are set, we can build azimuthally averaged radial profiles. The spaxels are binned into radius increments of 0.2 \reff\ from 0.0$-$2.6 \reff. While the MaNGA sample is limited spatially to 1.5$-$2.5 \reff\, the spaxels along the semi-minor axis of a galaxy will be able to extend beyond that range because of the galaxy's inclination angle, sin($i$).

@@

Figure~\ref{fig:nuc_sep} compares the enhancement in both SFR and sSFR as a function of projected separation. For our data points, we use the mean $\Delta\log{\rm SFR}$ and $\Delta\log{\rm sSFR}$ measured within a deprojected radius of 0.5~\reff\ (i.e., the black data points in the right side panel of Figure~\ref{fig:ssfr_sep}). The literature data is measured from single 1-1.5\arcsec-radius SDSS fibers \citep{Ellison:2008,Scudder:2012,Patton:2013,Bustamante:2020}. We find that our central SFR enhancements for closely separated pairs are higher than many of the previous studies, our sSFR enhancement is $\sim$0.2 dex higher than \citet{Scudder:2012} and $\sim$0.3 dex higher than \citet{Ellison:2008} and \citet{Bustamante:2020}. The central enhancement to the sSFR for closely separated pairs in our sample is $\sim$0.1 dex higher than \citet{Patton:2013}. The sSFR enhancement function from \citet{Li:2008} follows closely to our own and remains $\sim$0.05 dex above our sample out to 40 kpc.
